---
title: "Exploring marine mammals data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lwdata_cpod}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

<!--- Insert image
<img src='' class='rightimg'/>
-->

<!--- Describe datatype in one or two sentences. -->
Permanent Cetacean passive acoustic sensor network in the Belgian part of the North Sea

- Partners: [Flanders Marine Institute (VLIZ)](https://www.vliz.be/en/imis?module=institute&insid=36), [Royal Belgian Institute of Natural Sciences (RBINS)](https://www.vliz.be/en/imis?module=institute&insid=1883)
- Period: since January 2016
- Geographical coverage: [Belgian Exclusive Economic Zone](https://www.marineregions.org/gazetteer.php?p=details&id=3293)
- Taxonomic coverage: Habour porpoises (_[Phocoena phocoena](https://www.marinespecies.org/aphia.php?p=taxdetails&id=137117)_), other cetaceans, abiotic (ships sonar)
- Moratorium: None
- Data quality: Research-grade
 
## How to interpret this dataset
 
### Query options
* **Processing**: Type of data: one of Raw or Validated
* **Timeframe**: Starting and stopping date for the query.
* **Quality**:  Level of processing: default Hi+Mod.
* **Sample period**: Counts aggregated. One of 1 week, 1 day or 60 min. Also per minute when choosing Processing = Raw.
 
### Available columns
* **deployment_fk**: Link to the deployment metadata in the European Tracking Network database
* **Time**: Time in UTC, beginning of sample time bin.
* **Species**: NBHF = Harbour porpoises (Phocoena phocoena); other cet = other cetaceans; Sonar = sonar 
* **Milliseconds**: Duration of click trains in the specified time bin.
* **Number_clicks_filtered**: recognized clicks of a species.
* **Number_clicks_total**: total number of recorded clicks.
* **Lost_minutes**: Minutes in which the number of clicks was larger than 4096.
* **Recorded**: Minutes recording.
* **Dpm**: Detection positive minutes, or a mammal was detected in that minute. A click train of the selected species was recorded in that minute.
* **Dph**: Detection positive hours, or a mammal was detected in that hour. A click train of the selected species was recorded in that hour.
* **Dpd**: Detection positive days, or a mammal was detected that day. A click train of the selected species was recorded in that day.
* **Dpw**: Detection positive weeks, or a mammal was detected that day. A click train of the selected species was recorded in that week.
* **Quality**: RAW: 3=high; 2= mod; 1=low; Hi+mod=3; Hi+mod+low =3; Validated = hi+mod = 4.
* **Projectname**: Name of the project.
* **Station**: name or code of sampling station or receiver code.
* **Latitude**: Center latitude in decimal degrees, WGS84.
* **Longitude**: Center longitude in decimal degrees, WGS84.
* **Mooring_type**: Surface-buoy or bottom-mooring.
* **Receiver**: Receiver: serial number of C-POD.

## Abstract

```{r abstract, results = 'asis', echo = FALSE}
# Get the abstract and/or description from IMIS and print here
library(jsonlite)

dasid <- fromJSON("https://www.vliz.be/en/imis?module=dataset&dasid=5531&show=json")

abstract <- dasid[["datasetrec"]][["EngAbstract"]]
description <- dasid[["datasetrec"]][["EngDescr"]]

if (is.null(abstract) & is.null(description)){
  
} else if(is.null(abstract) & !is.null(description)){
  out <- description
} else if(!is.null(abstract) & is.null(description)){
  out <- abstract
} else if(!is.null(abstract) & !is.null(description)){
  out <- paste0(abstract, "</br>", description)
}

cat(out)
```

## How to cite these data

When using this tool, please acknowledge the LifeWatch Data Explorer in your work with the statment written below. Please also acknowledge the individual datasets used and mind their individual licenses: Find the complete list in this link to the [IMIS metadata record](https://www.vliz.be/en/imis?module=dataset&dasid=5531) of this data type.

> _This work makes use of the LifeWatch Data Explorer provided by VLIZ and funded by Research Foundation - Flanders (FWO) as part of the Belgian contribution to LifeWatch._

</br>
If you retrieved these data using the [LifeWatch Data Explorer R package](https://github.com/lifewatch/lwdataexplorer/), you should also cite it as: 

> ```{r citation, results = 'asis', echo = FALSE}
try(cat(citation("lwdataexplorer")$textVersion), TRUE)
```

