---
title: "Exploring Acoustic data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lwdata_acoustic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!--- Insert image  -->
<!-- <img src='' class='rightimg'/> -->

<!--- Describe datatype in one or two sentences. -->

Millidecade bands computed from the acoustic data collected in the LifeWatch Broadband Acoustic Network.
Frequency range from 0 to 24 kHz. Millidecade bands computed according to Miksis-Olds et al. (2021)<sup>1</sup>

- Partners: [Flanders Marine Institute (VLIZ)](https://www.vliz.be/en/imis?module=institute&insid=36), [Ghent University](https://www.vliz.be/en/imis?module=institute&insid=12685)
- Period: March 2021 - Ongoing
- Geographical coverage: [ANE, Belgium, Belgian Coast](http://marineregions.org/mrgid/2550), [ANE, Belgium, Belgian Continental Shelf (BCS)](http://marineregions.org/mrgid/3293)
- Taxonomical coverage: Animalia [WoRMS, 2023](https://www.marinespecies.org/aphia.php?p=taxdetails&id=2)
- Moratorium: None
- Data quality: proof-of-concept 

## How to interpret this dataset

### Query options
* **frequency**: Millidecade band (centre frequency)
* **Timeframe**: Starting and stopping date for the query.

### Available columns
* **frequency**: Centre frequency of the millidecade band, in Hz
* **upper_frequency**: Upper frequency limit of the millidecade band, in Hz
* **lower_frequency**: Lower frequency limit of the millidecade band, in Hz
* **value**: Sound pressure density of the millidecade band
* **datetime**: ISO 8601 Date and Time in UTC, beginning of sample time bin
* **Latitude**: Center latitude in decimal degrees, WGS84.
* **Longitude**: Center longitude in decimal degrees, WGS84.
* **station_name**: Name or code of sampling station.
* **hydrophone_depth**: Depth in the water column, in (positive) meters
* **etnID**: Unique identifier for each deployment, linked to the metadata specified in the European Tracking Network (https://www.lifewatch.be/etn/)

## Abstract
<!-- This information is pulled from the IMIS record of the data type -->

```{r abstract, results = 'asis', echo = FALSE}
# Get the abstract and/or description from IMIS and print here
library(jsonlite)

dasid <- fromJSON("https://www.vliz.be/en/imis?module=dataset&dasid=8172&show=json")

abstract <- dasid[["datasetrec"]][["EngAbstract"]]
description <- dasid[["datasetrec"]][["EngDescr"]]

if (is.null(abstract) & is.null(description)){

} else if(is.null(abstract) & !is.null(description)){
  out <- description
} else if(!is.null(abstract) & is.null(description)){
  out <- abstract
} else if(!is.null(abstract) & !is.null(description)){
  out <- paste0(abstract, "</br>", description)
}

cat(out)
```

## How to cite these data

```{r get citation and license, results = 'asis', echo = FALSE}
library(jsonlite)
# Set base url
url <- "https://www.vliz.be/en/imis?module=dataset&dasid=8172"
dasid <- fromJSON(paste0(url, "&show=json"))
# Build citation. Get active DOI.
citation <- dasid[["datasetrec"]][["Citation"]]
citation <- paste0("> ", citation, " ", url,
                   ". Accessed through the LifeWatch Data Explorer / lwdataexplorer R package.")
# Get license
license <- dasid[["datasetrec"]][["AccConstrDisplay"]]
```

```{r display citation, results = 'asis', echo = FALSE}
cat(citation)
```

```{r display license, results = 'asis', echo = FALSE}
cat(paste0("<b>Availability:</b>", license))
```

Please acknowledge as: This work makes use of the LifeWatch observation data and infrastructure (provided by VLIZ) funded by Research Foundation - Flanders (FWO) as part of the Belgian contribution to LifeWatch.

These data are also available in the R language with the [lwdataxplorer package](https://lifewatch.github.io/lwdataexplorer/).

<sup>1</sup><small>Miksis-Olds, J. L., Dugan, P. J., Martin, S. B., Klinck, H., Mellinger, D. K., Mann, D. A., Ponirakis, D. W., and Boebel, O. (2021). Ocean Sound Analysis Software for Making Ambient Noise Trends Accessible (MANTA). Frontiers in Marine Science, 8:703650. https://doi.org/10.3389/fmars.2021.703650 </small>
