---
title: "Exploring Sample Library"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lwdata_sample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
# <img src=''class='rightimg'/>
```

Query a sample library containing sediment, zooplankton, phytoplankton, eDNA and prokaryotic DNA samples.

- Partners: [European Marine Biological Resource Centre - Belgium (EMBRC)](https://www.vliz.be/en/imis?module=project&proid=4522), [LifeWatch Belgium](https://www.lifewatch.be)
- Period: since July 2012
- Geographical coverage: [Belgian Exclusive Economic Zone](https://www.marineregions.org/gazetteer.php?p=details&id=3293), [Southern bight of the North Sea](https://marineregions.org/gazetteer.php?p=details&id=2399)
- Taxonomic coverage: Macrobenthos, Zooplankton, Phytoplankton, Unknown Environmental DNA
- Moratorium: None
- Data quality: Unknown
 
## How to interpret this dataset
 
### Query options
* **Timeframe**: Starting and stopping date for the query.
 
### Available columns
* **Time**: Time in UTC, beginning of sample timeperiod.
* **Group**: Taxonomic group of organisms or group of artefacts.
* **Id**: Unique identifier to recognize a sample taken aboard the RV Simon Stevin.
* **TripActionId**: Unique identifier to recognize actions taken aboard the RV Simon Stevin.
* **Station**: Name or code of sampling station.
* **Latitude**: Center latitude in decimal degrees, WGS84.
* **Longitude**: Center longitude in decimal degrees, WGS84.
* **ActionType**: The type of sampling equipment. 
* **CurrentFixation**: Preservation method.
* **CurrentStorage**: Physical place where the sample is located.
* **Status**: One of: Triated, Scanned and validated, Collected, Not available, Filtered or Scanned.


## Abstract

Monthly LifeWatch sampling campaigns with the RV Simon Stevin and specific research surveys in the southern part of the North Sea contribute to the development of a marine biobank of the collected marine biological samples. This biobank has the objective to preserve and offer collected physical samples that can be used for biodiversity or ecosystem research upon request. A well-documented and accessible library of samples holds unique potential for retrospective research, e.g. providing insights on historical phenomena, assess baselines, trends, or simply exploit these historical samples with new techniques previously unavailable.

This LifeWatch explorer allows for a detailed search through the available samples. This facilitates a specific request for samples of interest based on spatial and temporal criteria.

## How to cite these data

Samples are collected, processed and preserved in the framework of the LifeWatch observatory, but made available as biological resources through [EMBRC.be](http://www.embrc.be). Therefore, **for use of the samples from the marine biobank, please provide a written request to [embrc@vliz.be](mailto:embrc@vliz.be)**.

When using this tool, please acknowledge the LifeWatch Data Explorer in your work with the statment written below.

> _This work makes use of the LifeWatch Data Explorer provided by VLIZ and funded by Research Foundation - Flanders (FWO) as part of the Belgian contribution to LifeWatch._

</br>
If you retrieved these data using the [LifeWatch Data Explorer R package](https://github.com/lifewatch/lwdataexplorer/), you should also cite it as: 

```{r citation, results = 'asis', echo = FALSE}
cat(paste0("> ", citation("lwdataexplorer")$textVersion))
```
